The Galactic Midplane Is Not a Plane: Implications for Dynamical Analysis with {\em Gaia} Data and Beyond

### 2019-05-04 paper/ms.tex
Fix simplified epicylic approx of Jphi
Move stuff about no error in Jphi earlier

Since most of the time stars will be at maximum amplitude (i.e. turnaround) in
both $R$ and $z$, we can approximate the order of magnitude of the systematic
error in the actions by
\beq\label{eq:Ji_err_mosttime}
\begin{split}
\frac{\Delta J_{\phi}}{J_{\phi}} &= \frac{\Delta R}{R_g} \\
\frac{\Delta J_{R}}{J_{R}} &= \frac{2\Delta R}{A_R} \\
\frac{\Delta J_{z}}{J_{z}} &= \frac{2\Delta z}{A_z} \text{,}
\end{split}
\eeq
where we have also ignored second order terms.

@@

For a systematic offset in $z$ ({\em upper} panels), the
95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $2.2$ and
$6.2\,\actunit$ for $J_R$ and $J_z$, respectively. These are fractional errors
of $5.7\,\%$ and $86\,\%$, respectively. The error induced in $J_{\phi}$ is
negligible, as expected since $J_{\phi}$ is independent of $z$. It is worth
pointing out that a $100\,\pc$ error in an orbit with
$z_{\text{max}}=850\,\pc$ --- a $12\%$ error
--- induced an $86\%$ error in the computation of $J_z$.

@@

In the {\em upper middle} panel of Figure~\ref{fig:many_orbit_wrong_ref},
there is essentially no spread in the determination of $J_{\phi}$. This is
expected since, in principle, $J_{\phi}$ only depends on the $x$- and
$y$-components of the position and velocity of the stars\footnote{In practice,
however, $J_{\phi}$ is computed as part of the torus-fitting method.}, and is
thus unaffected by offsets in $z$. Indeed, the result we found in
Figure~\ref{fig:one_orbit_wrong_ref} for the thick-disk orbit holds for all
orbit types. This is also a demonstration of the robustness of the orbit
integration and action calculation methods we use.

>>>>

Since most of the time stars will be at maximum amplitude (i.e. turnaround) in
both $R$ and $z$, we can approximate the order of magnitude of the systematic
error in the actions by
\beq\label{eq:Ji_err_mosttime}
\begin{split}
\frac{\Delta J_{\phi}}{J_{\phi}} &= \frac{\Delta R}{R_g} + \frac{\Delta v_{\phi}}{v_c} \\
\frac{\Delta J_{R}}{J_{R}} &= \frac{2\Delta R}{A_R} \\
\frac{\Delta J_{z}}{J_{z}} &= \frac{2\Delta z}{A_z} \text{,}
\end{split}
\eeq
where we have also ignored second order terms.

@@

For a systematic offset in $z$ ({\em upper} panels), the
95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $2.2$ and
$6.2\,\actunit$ for $J_R$ and $J_z$, respectively. These are fractional errors
of $5.7\,\%$ and $86\,\%$, respectively. The error induced in $J_{\phi}$ is
negligible, as expected since $J_{\phi}$ only depends on the $x$- and
$y$-components of the position and velocity of the stars.\footnote{In practice,
however, $J_{\phi}$ is computed as part of the torus-fitting method.} It is worth
pointing out that a $100\,\pc$ error in an orbit with
$z_{\text{max}}=850\,\pc$ --- a $12\%$ error
--- induced an $86\%$ error in the computation of $J_z$.

@@

In the {\em upper middle} panel of Figure~\ref{fig:many_orbit_wrong_ref},
there is essentially no spread in the determination of $J_{\phi}$. This is
expected since $J_{\phi}$ is independent of $z$ and is
thus unaffected by offsets in $z$, as discussed earlier. Indeed, the result we found in
Figure~\ref{fig:one_orbit_wrong_ref} for the thick-disk orbit holds for all
orbit types. This is also a demonstration of the robustness of the orbit
integration and action calculation methods we use.
